Hybrid platforms and bargaining power

### 2023-03-26 src/paper/application.tex
Add some proofs

For the following two propositions, let us ignore free entry and equilibrium behavior, and consider the number of fringe entrants $N_F$ as fixed.
The following statement says that the platform's profit is increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Assume that $f, w$ are continuously differentiable and that $w$ has a non-zero measure support.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    $$\frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0$$.
\end{proposition}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the size of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is also guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.
This decrease cannot be so strong, however, to decrease the platform's profits in absolute terms.
\cref{fig:increase_N_P_platform} shows an example of this situation.

@@

Next, let us look at an analogous result for the fringe firms.
As the next proposition demonstrates, in this case, the direction of the change depends on the complementarities between the fringe firms.
\begin{proposition}
    \label{prop:share_of_fringe}
    Assume that $f, w$ are twice continuously differentiable and that $w$ has a non-zero measure support.
    Let $N_F > 0$.
    Then $\pi_F$ is also differentiable.
    \begin{align*}
        &\text{If } \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} < 0 \;\forall n_F \leq N_F, \text{ then } \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} < 0, \\
        &\text{if } \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} > 0 \;\forall n_F \leq N_F, \text{ then } \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} > 0
    \end{align*}
    for all $N_P \geq 0$.
\end{proposition}
In summary, when they are mostly substitutes (the cross-derivatives of $f$ are negative), the fringe's profits decrease as a result of an increase in $N_F$.
The intuition is that, even though the total size of the pie increases, the bargaining power of the fringe deteriorates so much that its share decreases not only in relative, but also in absolute terms (as illustrated in \cref{fig:increase_N_P_fringe}).
On the other hand, when the fringe firms are mostly complements the fringe's profits increase.

@@

Now let us turn to equilibrium entry as a function of the platform's product variety.
The previous proposition implies an almost immediate corollary regarding the equilibrium number of fringe entrants.
\begin{corollary}
    \label{cor:fringe_entry}
    Assume that $f, w$ are twice continuously differentiable and that $w$ has a non-zero measure support.
    Let $N_F^*$ denote the equilibrium number of fringe firms.
    Let us also assume that $N_F^* > 0$.
    Then the equilibrium number of fringe firms is also differentiable and
    \begin{align*}
        &\text{If } \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0 \;\forall n_F \leq N_F, \text{ then } \frac{\partial N_F^*}{\partial N_P} < 0. \\
        &\text{If } \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} > 0 \;\forall n_F \leq N_F, \text{ then } \frac{\partial N_F^*}{\partial N_P} > 0.
    \end{align*}
\end{corollary}
That is, the equilibrium number of entrants increases as a response to an increase in $N_P$ if the fringe firms are mostly complements, and decreases if they are mostly substitutes.
The reason simply lies in the concave or hump-shaped fringe profit function.
If $N_F^*$ is an equilibrium, then fringe profits minus entry costs are strictly positive for all $N_F \leq N_F^*$ and strictly negative for all $N_F > N_F^*$.
Therefore, if an increase in $N_P$ decreases the fringe's profits for every $N_F \geq 0$, equilibrium can be restored by decreasing the number of fringe entrants, and vice versa for the other case (\cref{fig:comparative_N_F}).

@@

        \draw[->] (i-2 |- 0,0.5) -- (i2-2 |- 0,0.5);
    \end{tikzpicture}
    \caption{Comparative statistics for equilibrium entry (fringe are substitutes)}
    \label{fig:comparative_N_F}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
        \begin{axis}[
                xlabel={$N_P$},
                ylabel={$N_F^*$},
                xmin=0, xmax=1, % adjust these values as needed
                ymin=0, ymax=1, % adjust these values as needed
                % grid=major,
                % grid style={dashed,gray!30},
                width=6cm, % adjust the width of the plot
                height=6cm, % adjust the height of the plot
                axis lines=left,
                xtick=\empty, ytick=\empty
            ]
            \addplot[color=blue,mark=none,very thick] table[x=N_P,y=N_F]{\equilibrium};
            \addplot[color=black,mark=none,dashed] table[x=N_P,y=N_F_cf]{\equilibrium};
        \end{axis}
    \end{tikzpicture}
    \caption{The equilibrium number of fringe firms as a function of the number of platform products.}
    \label{fig:fringe_entry_eq}
\end{figure}

@@

>>>>

For the following two propositions, let us ignore free entry and equilibrium behavior, and consider the number of fringe entrants $N_F$ as fixed.
The following statement says that the platform's profit is increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    $$\frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0$$.
\end{proposition}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the size of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is also guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.
This decrease cannot be so strong, however, to decrease the platform's profits in absolute terms.
\cref{fig:increase_N_P_platform} shows an example of this situation.

@@

Next, let us look at an analogous result for the fringe firms.
As the next proposition demonstrates, in this case, the direction of the change depends on the complementarities between the fringe firms.
\begin{proposition}
    \label{prop:share_of_fringe}
    Assume that $f, w$ are twice continuously differentiable.
    Let $N_F > 0$.
    Then $\pi_F$ is also differentiable.
    \begin{align*}
        &\text{If } \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} < 0 \;\forall n_F \leq N_F, \text{ then } \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} < 0, \\
        &\text{if } \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} > 0 \;\forall n_F \leq N_F, \text{ then } \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} > 0
    \end{align*}
    for all $N_P \geq 0$.
\end{proposition}
In summary, when they are mostly substitutes (the cross-derivatives of $f$ are negative), the fringe's profits decrease as a result of an increase in $N_F$.
The intuition is that, even though the total size of the pie increases, the bargaining power of the fringe deteriorates so much that its share decreases not only in relative, but also in absolute terms (as illustrated in \cref{fig:increase_N_P_fringe}).
On the other hand, when the fringe firms are mostly complements the fringe's profits increase.

@@

Now let us turn to equilibrium entry as a function of the platform's product variety.
The previous proposition implies an almost immediate corollary regarding the equilibrium number of fringe entrants.
\begin{corollary}
    \label{cor:fringe_entry}
    Assume that $f, w$ are twice continuously differentiable.
    Let $N_F^*$ denote the equilibrium number of fringe firms.
    Let us also assume that $N_F^* > 0$.
    Furthermore, let us assume that
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0 \quad \forall n_F \leq N_F \\
        \frac{\partial^2 f(n_P, n_F)}{\partial n_F^2} < 0 \quad \forall n_F \leq N_F.
    \end{align*}
    Then the equilibrium number of fringe firms is also differentiable and
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < 0.
    \end{align*}
\end{corollary}
That is, the equilibrium number of entrants increases as a response to an increase in $N_P$ if the fringe firms are mostly complements, and decreases if they are mostly substitutes.
The reason simply lies in the concave or hump-shaped fringe profit function.
If $N_F^*$ is an equilibrium, then fringe profits minus entry costs are strictly positive for all $N_F \leq N_F^*$ and strictly negative for all $N_F > N_F^*$.
Therefore, if an increase in $N_P$ decreases the fringe's profits for every $N_F \geq 0$, equilibrium can be restored by decreasing the number of fringe entrants, and vice versa for the other case (\cref{fig:comparative_N_F}).

@@

        \draw[->] (i-2 |- 0,0.5) -- (i2-2 |- 0,0.5);
    \end{tikzpicture}
    \caption{Illustration of comparative statics for equilibrium entry}
    \label{fig:comparative_N_F}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{tikzpicture}
        \begin{axis}[
                xlabel={$N_P$},
                ylabel={$N_F^*$},
                xmin=0, xmax=1.5, % adjust these values as needed
                ymin=0, ymax=1.4, % adjust these values as needed
                width=10cm, % adjust the width of the plot
                height=6cm, % adjust the height of the plot
                axis lines=left,
                xtick=\empty, ytick=\empty
            ]
            \addplot[color=red,mark=none,very thick] table[x=N_P,y=N_F_bench]{\equilibrium};
            \addplot[color=blue,mark=none,very thick] table[x=N_P,y=N_F]{\equilibrium};
            \addplot[color=black,mark=none,thick,dashed] table[x=N_P,y=N_F_cf]{\equilibrium};
            \legend{Benchmark, Bargaining, Fixed aggregate}
        \end{axis}
    \end{tikzpicture}
    \caption{Equilibrium number of fringe firms as a function of the number of platform products.}
    \label{fig:fringe_entry_eq}
\end{figure}

@@

\section{Proofs and additional lemmas}

\begin{lemma}
    \label{lem:slope_at_eq}
    Let the conditions in \cref{ass:single_crossing} hold. Then for any $N_F^* > 0$ for which $\pi_F(N_P, N_F^*) = I_F$, the partial derivative with respect to the number of fringe firms is smaller than the investment cost:
    \begin{align*}
        \left. \frac{\partial \pi_F(N_P, N_F)}{\partial N_F} \right|_{N_F = N_F^*} < I_F.
    \end{align*}
\end{lemma}
\begin{proof}[Proof of \cref{lem:slope_at_eq}]
    Let us use Leibniz' rule to obtain the first two derivatives of the fringe's profit with respect to the number of fringe firms:
    \begin{align*}
        \frac{\partial f(N_P, sN_F)}{\partial N_F} &= \frac{\partial}{\partial N_F} \left[ f(N_P, N_F) - \int_0^1 w(s) f(N_P, sN_F) \ds \right] \\
        &= \frac{\partial f(N_P, N_F)}{\partial N_F} - \int_0^1 s w(s) \frac{\partial f(N_P, sN_F)}{\partial N_F} \ds, \\
        \frac{\partial^2 f(N_P, sN_F)}{\partial N_F^2} &= \frac{\partial^2}{\partial N_F^2} \left[ f(N_P, N_F) - \int_0^1 w(s) f(N_P, sN_F) \ds \right] \\
        &= \frac{\partial^2 f(N_P, N_F)}{\partial N_F^2} - \int_0^1 s^2 w(s) \frac{\partial f(N_P, sN_F)}{\partial N_F^2} \ds.
    \end{align*}
    Therefore, \cref{ass:single_crossing} implies that at any $N_F$, $\pi_F(N_P, N_F)$ is either concave or decreasing.

    In the remainder, for ease of notation, let us denote partial derivatives as follows:
    \begin{align*}
        \partial_F \pi_F(N_P^*, N_F^*) := \left. \frac{\partial \pi_F(N_P, N_F)}{\partial N_F} \right|_{N_P = N_P*, N_F = N_F^*}.
    \end{align*}

    Notice that the fact that $\pi_F$ is concave or decreasing in $N_F$ implies that if $\partial_F \pi_F(N_P, N_F')$ for some $N_F' \geq 0$, then $\partial_F \pi_F(N_P, N_F'') < \partial_F \pi_F(N_P, N_F')$ for any $\tilde{N_F} < \bar{N_F}$.

    Next, observe that if $\partial_F \pi_F(N_P, 0) < I_F$, then there is no equilibrium with $N_F > 0$.
    To show this, assume by contradiction that $\exists N_F^* > 0$ such that $\pi_F(N_P, N_F^*) = I_F N_F^*$.
    Then the mean value theorem implies that there is a $\bar{N}_F \in (0, N_F^*)$ such that $\partial_F \pi_F(N_P, \bar{N}_F) = I_F$.
    However, this clearly cannot be the case as $\partial_F \pi_F(N_P, N_F) < I_F$ for all $N_F > 0$.

    Now let $N_F^*$ be a positive number for which $f(N_P, N_F^*) = I_F N_F^*$, if it exists.
    It is easy to see that $\partial \pi(N_P, N_F^*) < I_F$.
    The reason is that if it exists, then $\partial_F \pi_F(N_P, 0) > I_F$.
    Now if $\partial_F \pi_F(N_P, N_F^*) \geq I_F \forall N_F > 0$, then $f(N_P, N_F) > I_F N_F$, and the two functions do not intersect.
    Therefore, there must exist some $\bar{N}_F > 0$ for which $\partial_F \pi_F(N_P, \bar{N}_F) < I_F$.
    This in turn implies that $\partial_F \pi_F(N_P, N_F) \leq \partial_F \pi_F(N_P, \bar{N}_F) < I_F$ for all $N_F > \bar{N}_F$.
\end{proof}

\begin{proof}[Proof of \cref{prop:unique_equilibrium}]
    \Cref{lem:slope_at_eq} states that for any positive $N_F^*$ for which $\pi_F(N_P, N_F^*)$, the partial derivative with respect to the number of fringe firms is negative.
    Now assume by contradiction that $\exists 0 < N_F^* < N_F^{**}$ such that $\pi_F(N_P, N_F^*) = I_F N_F^*$ and $\pi_F(N_P, N_F^{**}) = I_F N_F^{**}$.
    But then the mean value theorem implies that there is a $\bar{N}_F \in (N_F^*, N_F^{**})$ such that $\partial_F \pi_F(N_P, \bar{N}_F) = I_F$.
    This is a contradiction, as $\partial_F \pi_F(N_P, N_F) < I_F$ for all $N_F^* < N_F$.
\end{proof}

\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} &= \frac{\partial}{\partial N_P} \int_0^1 w(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0.
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{prop:share_of_fringe}]
    Remember that
    \begin{align*}
        \pi_F(N_P, N_F) = f(N_P, N_F) - \int_0^1 w(s) f(N_P, N_F) \ds.
    \end{align*}
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} &= \frac{\partial}{\partial N_P} \left[ f(N_P, N_F) - \int_0^1 w(s) f(N_P, N_F) \ds \right]\\
        &= \frac{\partial f(N_P, N_F)}{\partial N_P} - \int_0^1 w(s) \frac{\partial f(N_P, sN_F)}{\partial N_P} \ds \\
        &= \underbrace{h(N_F)}_{(*)} - \underbrace{\int_0^1 w(s) h(s N_F)}_{(**)}
    \end{align*}
    where $h(N_F) = \frac{\partial h(N_F)}{\partial N_P}$.

    Notice that $(*)$ and $(**)$ are both positive.
    Furthermore, $(**)$ is the weighed average of $(*)$ over the interval $[0, N_F]$.
    Therefore, if $h(N_F)$ is strictly increasing, then $(*) > (**)$, and the whole expression is positive.
    Conversely, it is negative if $h(N_F)$ is strictly decreasing.

    If $h(N_F)$ is differentiable, then it is equivalent to saying that
    \begin{align*}
        \frac{\mathrm{d}h(N_F)}{\mathrm{d}N_F} = \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} < 0 &\implies \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} < 0 \\
        \frac{\mathrm{d}h(N_F)}{\mathrm{d}N_F} = \frac{\partial^2 f(N_P, n_F)}{\partial n_P \partial n_F} > 0 &\implies \frac{\partial \pi_F(N_P, N_F)}{\partial N_P} > 0
    \end{align*}
\end{proof}

\begin{proof}[Proof of \cref{cor:fringe_entry}]
    \Cref{prop:unique_equilibrium} establishes that if an equilibrium with $N_F > 0$ exists, it is unique.
    Use the implicit function theorem on the equation from \cref{ass:free_entry} to obtain
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} = \frac{\frac{\partial \pi_F(N_P, N_F)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F)}{\partial N_F}}
    \end{align*}

    Remember from \cref{prop:share_of_fringe} that the numerator of this expression is negative under the condition $\frac{\partial^2 f(N_P, N_F)}{\partial N_P \partial N_F}$.
    All we need to show to conclude the proof is that
    \begin{align*}
        \frac{\partial \pi_F (N_P, N_F)}{\partial N_F} < I_F,
    \end{align*}
    which is the statement of \cref{lem:slope_at_eq}.
\end{proof}
